---
title: "Looking at what matters the most"
author: "Maurizio La Rosa"
date: "14/11/2020"
output: 
  html_document: 
    theme: flatly
    highlight: pygments
    df_print: paged
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,fig.dim=c(5,5))
options(scipen=999)
setwd("/Users/mau/github/covid_19_data")
```

```{r, echo=FALSE}
options(scipen=999)
setwd("/Users/mau/github/covid_19_data")
covid_data <- read.delim("covid_data.csv",sep=",",header=T)
date_matrix <- matrix(covid_data$date,ncol=22)
date <- as.Date(date_matrix[,1])
tamponi_matrix <- matrix(covid_data$tamponi,ncol=22)
tamponi_ita <- diff(tamponi_matrix[,1])
```

# The problem of the data

The most problematic issue about the way the Italian government has decided to present the data about the evolution of the pandemic is that the data themselves were (and to a great extent still are) not accurate. This fact has mislead the perception of the health emergency by the public and has made it difficult to develop a rational way of fighting it, even on the account of the institutions. It alsa has determined that very different situations have been treated as if they were the same. The problem of the (missing) accuracy of the data relates to different aspects of the data collection. While I try to show the most obvious, I will certainly miss arguments that are less obvious:

- It is fairly clear that the most effective way of saying if a person has been infected is to test her. Although straightforward, this simple fact clashes with the harsh reality that there can never be enough tests to be performed daily by the healthcare system. This has been true during the first wave, when Italy had not enough tests available, but it remains true today, simply because the healthcare system cannot afford to perform a large amount of tests, in any case. What has greatly changed between the first and the second wave is the capacity to test: this has been greatly increased 

```{r, echo=FALSE}
#Function that allows me to customize the plot region
plot_region <- function() {
  points(0, 0, pch=16, cex=1e6, col="#FFFFFF") #FFFFFF: white
  #grid(NA,ny=NULL,col="#EAF2F3",lty=1,lwd=2)
}
```

```{r, echo=FALSE}
ymarks <- c(0,10000,40000,80000,120000,160000,200000,240000,280000)
ylabels <- c("0","10","40","80","120","160","200","240","280")
xmarks <- date[c(1,21,82,143,205,266,320)] #,"2021-01-15"
xlabels <- c("24 Feb","15 Mar","15 Mag","15 Lug","15 Set","15 Nov","8 Gen")

par(bg="#EAF2F3")
plot(c(0,tamponi_ita)~date,ylim=c(0,max(tamponi_ita)),type="l",xaxt="n",yaxt="n",xlab="",ylab="",bty="n",main="Tamponi giornalieri")
title(xlab="Mese",ylab="Numero di tamponi (migliaia)",cex.lab=.9,col.lab="#696969",panel.first=plot_region())
axis.Date(1,at=xmarks,labels=xlabels,cex.axis=.8,col="#696969",col.axis="#696969",las=2)
axis(2,tick=TRUE,at=ymarks,labels=ylabels,las=2,cex.axis=.8,col="#696969",col.axis="#696969")
abline(h=ymarks,col="#EAF2F3",panel.first=plot_region()) #v=xmarks,
points(c(0,tamponi_ita)~date,type="l",col="#8877AACC")
mean_tamp <- data.frame("Mean first wave"=round(mean(tamponi_ita[1:30])),"Mean second wave"=round(mean(tamponi_ita[214:244])))
knitr::kable(mean_tamp,caption="Mean daily tests in the first thirty days of the two waves")
summary(diff(tamponi_matrix)[15,2:22])
hist(diff(tamponi_matrix)[15,2:22],panel.first=plot_region())
means_tamp_reg_fw <- colSums(diff(tamponi_matrix)[1:29,])/29
means_fw <- mean(means_tamp_reg_fw[2:22])
sd_fw <- sd(means_tamp_reg_fw[2:22])
means_tamp_reg_sw <- colSums(diff(tamponi_matrix)[214:243,])/29
means_sw <- mean(means_tamp_reg_sw[2:22])
sd_sw <- sd(means_tamp_reg_sw[2:22])
var_tamp <- data.frame("First wave"=round(c(means_fw,sd_fw)),"Second wave"=round(c(means_sw,sd_sw)))
knitr::kable(var_tamp,caption="Variability in mean daily tests in the first thirty days of the two waves")
```

```{r, echo=FALSE}
#m_totale_casi <- matrix(covid_data$totale_casi,ncol=22)
#m_primi_casi=matrix(covid_data$nuovi_positivi,ncol=22)[1,]
#m_nuovi_casi <- rbind(m_primi_casi,m_totale_casi)
m_tamponi <- matrix(covid_data$tamponi,ncol=22)
m_primi_tamponi=matrix(covid_data$tamponi,ncol=22)[1,]
m_nuovi_tamponi <- rbind(m_primi_tamponi,diff(m_tamponi))
quota_positivi <- covid_data$nuovi_positivi/m_nuovi_tamponi

matrix(c(covid_data$nuovi_positivi),m_totale_casi)
quota_positivi_nord <- rowSums(matrix(c(covid_data$nuovi_positivi[1],diff(covid_data$totale_casi))/c(covid_data$tamponi[1],diff(covid_data$tamponi)),ncol=22)[,c(2:8,21,22)])
quota_positivi_centro <- rowSums(matrix(c(covid_data$nuovi_positivi[1],diff(covid_data$totale_casi))/c(covid_data$tamponi[1],diff(covid_data$tamponi)),ncol=22)[,9:12])
quota_positivi_sud <- rowSums(matrix(c(covid_data$nuovi_positivi[1],diff(covid_data$totale_casi))/c(covid_data$tamponi[1],diff(covid_data$tamponi)),ncol=22)[,13:20])

plot(quota_positivi_nord,type="l",col="blue")
lines(quota_positivi_centro,col="green")
lines(quota_positivi_sud,col="orange")

terapia_int_nord <- rowSums(matrix(covid_data$terapia_intensiva,ncol=22)[,c(2:8,21,22)])
terapia_int_centro <- rowSums(matrix(covid_data$terapia_intensiva,ncol=22)[,9:12])
terapia_int_sud <- rowSums(matrix(covid_data$terapia_intensiva,ncol=22)[,13:20])

plot(terapia_int_nord,type="l",col="blue")
lines(terapia_int_centro,col="green")
lines(terapia_int_sud,col="orange")

deceduti_nord <- rowSums(matrix(covid_data$deceduti,ncol=22)[,c(2:8,21,22)])
deceduti_centro <- rowSums(matrix(covid_data$deceduti,ncol=22)[,9:12])
deceduti_sud <- rowSums(matrix(covid_data$deceduti,ncol=22)[,13:20])

guariti_nord <- rowSums(matrix(covid_data$dimessi_guariti,ncol=22)[,c(2:8,21,22)])
guariti_centro <- rowSums(matrix(covid_data$dimessi_guariti,ncol=22)[,9:12])
guariti_sud <- rowSums(matrix(covid_data$dimessi_guariti,ncol=22)[,13:20])
```

regioni_nord <- data.frame(rbind(
  subset(covid_data,covid_data$codice_regione==1),
  subset(covid_data,covid_data$codice_regione==2),
  subset(covid_data,covid_data$codice_regione==3),
  subset(covid_data,covid_data$codice_regione==5),
  subset(covid_data,covid_data$codice_regione==6),
  subset(covid_data,covid_data$codice_regione==7),
  subset(covid_data,covid_data$codice_regione==8),
  subset(covid_data,covid_data$codice_regione==21),
  subset(covid_data,covid_data$codice_regione==22)))
regioni_centro <- data.frame(rbind(
  subset(covid_data,covid_data$codice_regione==9),
  subset(covid_data,covid_data$codice_regione==10),
  subset(covid_data,covid_data$codice_regione==11),
  subset(covid_data,covid_data$codice_regione==12)))
regioni_sud <- data.frame(rbind(
  subset(covid_data,covid_data$codice_regione==13),
  subset(covid_data,covid_data$codice_regione==14),
  subset(covid_data,covid_data$codice_regione==15),
  subset(covid_data,covid_data$codice_regione==16),
  subset(covid_data,covid_data$codice_regione==17),
  subset(covid_data,covid_data$codice_regione==18),
  subset(covid_data,covid_data$codice_regione==19),
  subset(covid_data,covid_data$codice_regione==20)))
```

```{r, echo=FALSE}
###This code chunk allows me to load the data and treat dates conveniently
#setwd("github/covid_19_data/")
covid_data <- read.delim("covid_data.csv",sep=",",header=T)
covid_data$date <- as.Date(covid_data$date)
regions <- levels(covid_data$denominazione_regione)

###This code chunk allows me to subset the Italian regions.
italia <- subset(covid_data,covid_data$denominazione_regione=="Italia")
n <- length(italia$date) #convenient, because it represents the number of observations
italia$growth_tot <- c(0,diff(italia$totale_casi)/italia$totale_casi[1:n-1]) #Growth rate of total cases
italia$nuovi_terapia_intensiva <- c(0,diff(italia$terapia_intensiva))
italia$growth_ter_int <- c(0,diff(italia$terapia_intensiva)/italia$terapia_intensiva[1:n-1]) #Growth rate of intensive care
italia$nuovi_deceduti <- c(0,diff(italia$deceduti))
italia$growth_dec <- c(0,diff(italia$deceduti)/italia$deceduti[1:n-1]) #Growth rate of deceased
italia$nuovi_guariti <- c(0,diff(italia$dimessi_guariti))

abruzzo <- subset(covid_data,covid_data$denominazione_regione=="Abruzzo")
abruzzo$growth_tot <- c(0,diff(abruzzo$totale_casi)/abruzzo$totale_casi[1:n-1]) #Growth rate of contagions
abruzzo$growth_ter_int <- c(0,diff(abruzzo$terapia_intensiva)/abruzzo$terapia_intensiva[1:n-1]) #Growth rate of intensive care
abruzzo$nuovi_deceduti <- c(0,diff(abruzzo$deceduti))
abruzzo$growth_dec <- c(0,diff(abruzzo$deceduti)/abruzzo$deceduti[1:n-1]) #Growth rate of deceased
abruzzo$share_tot <- abruzzo$totale_casi/italia$totale_casi
abruzzo$share_new <- abruzzo$nuovi_positivi/italia$nuovi_positivi
abruzzo$share_ter_int <- abruzzo$terapia_intensiva/italia$terapia_intensiva
abruzzo$share_dec <- abruzzo$deceduti/italia$deceduti

basilicata <- subset(covid_data,covid_data$denominazione_regione=="Basilicata")
basilicata$growth_tot <- c(0,diff(basilicata$totale_casi)/basilicata$totale_casi[1:n-1]) #Growth rate of contagions
basilicata$growth_ter_int <- c(0,diff(basilicata$terapia_intensiva)/basilicata$terapia_intensiva[1:n-1]) #Growth rate of intensive care
basilicata$nuovi_deceduti <- c(0,diff(basilicata$deceduti))
basilicata$growth_dec <- c(0,diff(basilicata$deceduti)/basilicata$deceduti[1:n-1]) #Growth rate of deceased
basilicata$share_tot <- basilicata$totale_casi/italia$totale_casi
basilicata$share_new <- basilicata$nuovi_positivi/italia$nuovi_positivi
basilicata$share_ter_int <- basilicata$terapia_intensiva/italia$terapia_intensiva
basilicata$share_dec <- basilicata$deceduti/italia$deceduti

calabria <- subset(covid_data,covid_data$denominazione_regione=="Calabria")
calabria$growth_tot <- c(0,diff(calabria$totale_casi)/calabria$totale_casi[1:n-1]) #Growth rate of contagions
calabria$growth_ter_int <- c(0,diff(calabria$terapia_intensiva)/calabria$terapia_intensiva[1:n-1]) #Growth rate of intensive care
calabria$nuovi_deceduti <- c(0,diff(calabria$deceduti))
calabria$growth_dec <- c(0,diff(calabria$deceduti)/calabria$deceduti[1:n-1]) #Growth rate of deceased
calabria$share_tot <- calabria$totale_casi/italia$totale_casi
calabria$share_new <- calabria$nuovi_positivi/italia$nuovi_positivi
calabria$share_ter_int <- calabria$terapia_intensiva/italia$terapia_intensiva
calabria$share_dec <- calabria$deceduti/italia$deceduti

campania <- subset(covid_data,covid_data$denominazione_regione=="Campania")
campania$growth_tot <- c(0,diff(campania$totale_casi)/campania$totale_casi[1:n-1]) #Growth rate of contagions
campania$growth_ter_int <- c(0,diff(campania$terapia_intensiva)/campania$terapia_intensiva[1:n-1]) #Growth rate of intensive care
campania$nuovi_deceduti <- c(0,diff(campania$deceduti))
campania$growth_dec <- c(0,diff(campania$deceduti)/campania$deceduti[1:n-1]) #Growth rate of deceased
campania$share_tot <- campania$totale_casi/italia$totale_casi
campania$share_new <- campania$nuovi_positivi/italia$nuovi_positivi
campania$share_ter_int <- campania$terapia_intensiva/italia$terapia_intensiva
campania$share_dec <- campania$deceduti/italia$deceduti

e_romagna <- subset(covid_data,covid_data$denominazione_regione=="Emilia-Romagna")
e_romagna$growth_tot <- c(0,diff(e_romagna$totale_casi)/e_romagna$totale_casi[1:n-1]) #Growth rate of contagions
e_romagna$growth_ter_int <- c(0,diff(e_romagna$terapia_intensiva)/e_romagna$terapia_intensiva[1:n-1]) #Growth rate of intensive care
e_romagna$nuovi_deceduti <- c(0,diff(e_romagna$deceduti))
e_romagna$growth_dec <- c(0,diff(e_romagna$deceduti)/e_romagna$deceduti[1:n-1]) #Growth rate of deceased
e_romagna$share_tot <- e_romagna$totale_casi/italia$totale_casi
e_romagna$share_new <- e_romagna$nuovi_positivi/italia$nuovi_positivi
e_romagna$share_ter_int <- e_romagna$terapia_intensiva/italia$terapia_intensiva
e_romagna$share_dec <- e_romagna$deceduti/italia$deceduti

fv_giulia <- subset(covid_data,covid_data$denominazione_regione=="Friuli Venezia Giulia")
fv_giulia$growth_tot <- c(0,diff(fv_giulia$totale_casi)/fv_giulia$totale_casi[1:n-1]) #Growth rate of contagions
fv_giulia$growth_ter_int <- c(0,diff(fv_giulia$terapia_intensiva)/fv_giulia$terapia_intensiva[1:n-1]) #Growth rate of intensive care
fv_giulia$nuovi_deceduti <- c(0,diff(fv_giulia$deceduti))
fv_giulia$growth_dec <- c(0,diff(fv_giulia$deceduti)/fv_giulia$deceduti[1:n-1]) #Growth rate of deceased
fv_giulia$share_tot <- fv_giulia$totale_casi/italia$totale_casi
fv_giulia$share_new <- fv_giulia$nuovi_positivi/italia$nuovi_positivi
fv_giulia$share_ter_int <- fv_giulia$terapia_intensiva/italia$terapia_intensiva
fv_giulia$share_dec <- fv_giulia$deceduti/italia$deceduti

lazio <- subset(covid_data,covid_data$denominazione_regione=="Lazio")
lazio$growth_tot <- c(0,diff(lazio$totale_casi)/lazio$totale_casi[1:n-1]) #Growth rate of contagions
lazio$growth_ter_int <- c(0,diff(lazio$terapia_intensiva)/lazio$terapia_intensiva[1:n-1]) #Growth rate of intensive care
lazio$nuovi_deceduti <- c(0,diff(lazio$deceduti))
lazio$growth_dec <- c(0,diff(lazio$deceduti)/lazio$deceduti[1:n-1]) #Growth rate of deceased
lazio$share_tot <- lazio$totale_casi/italia$totale_casi
lazio$share_new <- lazio$nuovi_positivi/italia$nuovi_positivi
lazio$share_ter_int <- lazio$terapia_intensiva/italia$terapia_intensiva
lazio$share_dec <- lazio$deceduti/italia$deceduti

liguria <- subset(covid_data,covid_data$denominazione_regione=="Liguria")
liguria$growth_tot <- c(0,diff(liguria$totale_casi)/liguria$totale_casi[1:n-1]) #Growth rate of contagions
liguria$growth_ter_int <- c(0,diff(liguria$terapia_intensiva)/liguria$terapia_intensiva[1:n-1]) #Growth rate of intensive care
liguria$nuovi_deceduti <- c(0,diff(liguria$deceduti))
liguria$growth_dec <- c(0,diff(liguria$deceduti)/liguria$deceduti[1:n-1]) #Growth rate of deceased
liguria$share_tot <- liguria$totale_casi/italia$totale_casi
liguria$share_new <- liguria$nuovi_positivi/italia$nuovi_positivi
liguria$share_ter_int <- liguria$terapia_intensiva/italia$terapia_intensiva
liguria$share_dec <- liguria$deceduti/italia$deceduti

lombardia <- subset(covid_data,covid_data$denominazione_regione=="Lombardia")
lombardia$growth_tot <- c(0,diff(lombardia$totale_casi)/lombardia$totale_casi[1:n-1]) #Growth rate of contagions
lombardia$growth_ter_int <- c(0,diff(lombardia$terapia_intensiva)/lombardia$terapia_intensiva[1:n-1]) #Growth rate of intensive care
lombardia$nuovi_deceduti <- c(0,diff(lombardia$deceduti))
lombardia$growth_dec <- c(0,diff(lombardia$deceduti)/lombardia$deceduti[1:n-1]) #Growth rate of deceased
lombardia$share_tot <- lombardia$totale_casi/italia$totale_casi
lombardia$share_new <- lombardia$nuovi_positivi/italia$nuovi_positivi
lombardia$share_ter_int <- lombardia$terapia_intensiva/italia$terapia_intensiva
lombardia$share_dec <- lombardia$deceduti/italia$deceduti

marche <- subset(covid_data,covid_data$denominazione_regione=="Marche")
marche$growth_tot <- c(0,diff(marche$totale_casi)/marche$totale_casi[1:n-1]) #Growth rate of contagions
marche$growth_ter_int <- c(0,diff(marche$terapia_intensiva)/marche$terapia_intensiva[1:n-1]) #Growth rate of intensive care
marche$nuovi_deceduti <- c(0,diff(marche$deceduti))
marche$growth_dec <- c(0,diff(marche$deceduti)/marche$deceduti[1:n-1]) #Growth rate of deceased
marche$share_tot <- marche$totale_casi/italia$totale_casi
marche$share_new <- marche$nuovi_positivi/italia$nuovi_positivi
marche$share_ter_int <- marche$terapia_intensiva/italia$terapia_intensiva
marche$share_dec <- marche$deceduti/italia$deceduti

molise <- subset(covid_data,covid_data$denominazione_regione=="Molise")
molise$growth_tot <- c(0,diff(molise$totale_casi)/molise$totale_casi[1:n-1]) #Growth rate of contagions
molise$growth_ter_int <- c(0,diff(molise$terapia_intensiva)/molise$terapia_intensiva[1:n-1]) #Growth rate of intensive care
molise$nuovi_deceduti <- c(0,diff(molise$deceduti))
molise$growth_dec <- c(0,diff(molise$deceduti)/molise$deceduti[1:n-1]) #Growth rate of deceased
molise$share_tot <- molise$totale_casi/italia$totale_casi
molise$share_new <- molise$nuovi_positivi/italia$nuovi_positivi
molise$share_ter_int <- molise$terapia_intensiva/italia$terapia_intensiva
molise$share_dec <- molise$deceduti/italia$deceduti

bolzano <- subset(covid_data,covid_data$denominazione_regione=="P.A. Bolzano")
bolzano$growth_tot <- c(0,diff(bolzano$totale_casi)/bolzano$totale_casi[1:n-1]) #Growth rate of contagions
bolzano$growth_ter_int <- c(0,diff(bolzano$terapia_intensiva)/bolzano$terapia_intensiva[1:n-1]) #Growth rate of intensive care
bolzano$nuovi_deceduti <- c(0,diff(bolzano$deceduti))
bolzano$growth_dec <- c(0,diff(bolzano$deceduti)/bolzano$deceduti[1:n-1]) #Growth rate of deceased
bolzano$share_tot <- bolzano$totale_casi/italia$totale_casi
bolzano$share_new <- bolzano$nuovi_positivi/italia$nuovi_positivi
bolzano$share_ter_int <- bolzano$terapia_intensiva/italia$terapia_intensiva
bolzano$share_dec <- bolzano$deceduti/italia$deceduti

trento <- subset(covid_data,covid_data$denominazione_regione=="P.A. Trento")
trento$growth_tot <- c(0,diff(trento$totale_casi)/trento$totale_casi[1:n-1]) #Growth rate of contagions
trento$growth_ter_int <- c(0,diff(trento$terapia_intensiva)/trento$terapia_intensiva[1:n-1]) #Growth rate of intensive care
trento$nuovi_deceduti <- c(0,diff(trento$deceduti))
trento$growth_dec <- c(0,diff(trento$deceduti)/trento$deceduti[1:n-1]) #Growth rate of deceased
trento$share_tot <- trento$totale_casi/italia$totale_casi
trento$share_new <- trento$nuovi_positivi/italia$nuovi_positivi
trento$share_ter_int <- trento$terapia_intensiva/italia$terapia_intensiva
trento$share_dec <- trento$deceduti/italia$deceduti

piemonte <- subset(covid_data,covid_data$denominazione_regione=="Piemonte")
piemonte$growth_tot <- c(0,diff(piemonte$totale_casi)/piemonte$totale_casi[1:n-1]) #Growth rate of contagions
piemonte$growth_ter_int <- c(0,diff(piemonte$terapia_intensiva)/piemonte$terapia_intensiva[1:n-1]) #Growth rate of intensive care
piemonte$nuovi_deceduti <- c(0,diff(piemonte$deceduti))
piemonte$growth_dec <- c(0,diff(piemonte$deceduti)/piemonte$deceduti[1:n-1]) #Growth rate of deceased
piemonte$share_tot <- piemonte$totale_casi/italia$totale_casi
piemonte$share_new <- piemonte$nuovi_positivi/italia$nuovi_positivi
piemonte$share_ter_int <- piemonte$terapia_intensiva/italia$terapia_intensiva
piemonte$share_dec <- piemonte$deceduti/italia$deceduti

puglia <- subset(covid_data,covid_data$denominazione_regione=="Puglia")
puglia$growth_tot <- c(0,diff(puglia$totale_casi)/puglia$totale_casi[1:n-1]) #Growth rate of contagions
puglia$growth_ter_int <- c(0,diff(puglia$terapia_intensiva)/puglia$terapia_intensiva[1:n-1]) #Growth rate of intensive care
puglia$nuovi_deceduti <- c(0,diff(puglia$deceduti))
puglia$growth_dec <- c(0,diff(puglia$deceduti)/puglia$deceduti[1:n-1]) #Growth rate of deceased
puglia$share_tot <- puglia$totale_casi/italia$totale_casi
puglia$share_new <- puglia$nuovi_positivi/italia$nuovi_positivi
puglia$share_ter_int <- puglia$terapia_intensiva/italia$terapia_intensiva
puglia$share_dec <- puglia$deceduti/italia$deceduti

sardegna <- subset(covid_data,covid_data$denominazione_regione=="Sardegna")
sardegna$growth_tot <- c(0,diff(sardegna$totale_casi)/sardegna$totale_casi[1:n-1]) #Growth rate of contagions
sardegna$growth_ter_int <- c(0,diff(sardegna$terapia_intensiva)/sardegna$terapia_intensiva[1:n-1]) #Growth rate of intensive care
sardegna$nuovi_deceduti <- c(0,diff(sardegna$deceduti))
sardegna$growth_dec <- c(0,diff(sardegna$deceduti)/sardegna$deceduti[1:n-1]) #Growth rate of deceased
sardegna$share_tot <- sardegna$totale_casi/italia$totale_casi
sardegna$share_new <- sardegna$nuovi_positivi/italia$nuovi_positivi
sardegna$share_ter_int <- sardegna$terapia_intensiva/italia$terapia_intensiva
sardegna$share_dec <- sardegna$deceduti/italia$deceduti

sicilia <- subset(covid_data,covid_data$denominazione_regione=="Sicilia")
sicilia$growth_tot <- c(0,diff(sicilia$totale_casi)/sicilia$totale_casi[1:n-1]) #Growth rate of contagions
sicilia$growth_ter_int <- c(0,diff(sicilia$terapia_intensiva)/sicilia$terapia_intensiva[1:n-1]) #Growth rate of intensive care
sicilia$nuovi_deceduti <- c(0,diff(sicilia$deceduti))
sicilia$growth_dec <- c(0,diff(sicilia$deceduti)/sicilia$deceduti[1:n-1]) #Growth rate of deceased
sicilia$share_tot <- sicilia$totale_casi/italia$totale_casi
sicilia$share_new <- sicilia$nuovi_positivi/italia$nuovi_positivi
sicilia$share_ter_int <- sicilia$terapia_intensiva/italia$terapia_intensiva
sicilia$share_dec <- sicilia$deceduti/italia$deceduti

toscana <- subset(covid_data,covid_data$denominazione_regione=="Toscana")
toscana$growth_tot <- c(0,diff(toscana$totale_casi)/toscana$totale_casi[1:n-1]) #Growth rate of contagions
toscana$growth_ter_int <- c(0,diff(toscana$terapia_intensiva)/toscana$terapia_intensiva[1:n-1]) #Growth rate of intensive care
toscana$nuovi_deceduti <- c(0,diff(toscana$deceduti))
toscana$growth_dec <- c(0,diff(toscana$deceduti)/toscana$deceduti[1:n-1]) #Growth rate of deceased
toscana$share_tot <- toscana$totale_casi/italia$totale_casi
toscana$share_new <- toscana$nuovi_positivi/italia$nuovi_positivi
toscana$share_ter_int <- toscana$terapia_intensiva/italia$terapia_intensiva
toscana$share_dec <- toscana$deceduti/italia$deceduti

umbria <- subset(covid_data,covid_data$denominazione_regione=="Umbria")
umbria$growth_tot <- c(0,diff(umbria$totale_casi)/umbria$totale_casi[1:n-1]) #Growth rate of contagions
umbria$growth_ter_int <- c(0,diff(umbria$terapia_intensiva)/umbria$terapia_intensiva[1:n-1]) #Growth rate of intensive care
umbria$nuovi_deceduti <- c(0,diff(umbria$deceduti))
umbria$growth_dec <- c(0,diff(umbria$deceduti)/umbria$deceduti[1:n-1]) #Growth rate of deceased
umbria$share_tot <- umbria$totale_casi/italia$totale_casi
umbria$share_new <- umbria$nuovi_positivi/italia$nuovi_positivi
umbria$share_ter_int <- umbria$terapia_intensiva/italia$terapia_intensiva
umbria$share_dec <- umbria$deceduti/italia$deceduti

v_aosta <- subset(covid_data,covid_data$denominazione_regione=="Valle d'Aosta")
v_aosta$growth_tot <- c(0,diff(v_aosta$totale_casi)/v_aosta$totale_casi[1:n-1]) #Growth rate of contagions
v_aosta$growth_ter_int <- c(0,diff(v_aosta$terapia_intensiva)/v_aosta$terapia_intensiva[1:n-1]) #Growth rate of intensive care
v_aosta$nuovi_deceduti <- c(0,diff(v_aosta$deceduti))
v_aosta$growth_dec <- c(0,diff(v_aosta$deceduti)/v_aosta$deceduti[1:n-1]) #Growth rate of deceased
v_aosta$share_tot <- v_aosta$totale_casi/italia$totale_casi
v_aosta$share_new <- v_aosta$nuovi_positivi/italia$nuovi_positivi
v_aosta$share_ter_int <- v_aosta$terapia_intensiva/italia$terapia_intensiva
v_aosta$share_dec <- v_aosta$deceduti/italia$deceduti

veneto <- subset(covid_data,covid_data$denominazione_regione=="Veneto")
veneto$growth_tot <- c(0,diff(veneto$totale_casi)/veneto$totale_casi[1:n-1]) #Growth rate of contagions
veneto$growth_ter_int <- c(0,diff(veneto$terapia_intensiva)/veneto$terapia_intensiva[1:n-1]) #Growth rate of intensive care
veneto$nuovi_deceduti <- c(0,diff(veneto$deceduti))
veneto$growth_dec <- c(0,diff(veneto$deceduti)/veneto$deceduti[1:n-1]) #Growth rate of deceased
veneto$share_tot <- veneto$totale_casi/italia$totale_casi
veneto$share_new <- veneto$nuovi_positivi/italia$nuovi_positivi
veneto$share_ter_int <- veneto$terapia_intensiva/italia$terapia_intensiva
veneto$share_dec <- veneto$deceduti/italia$deceduti
############################
############################
############################

#Latest observation of contagions (reproducing the table from Ministero della Salute: )
Data <- as.vector(italia$date)[length(italia$date)]
Regione <- as.vector(append(regions,"Italia"))
latest_abruzzo <- abruzzo[n,c(4,11,14,15,18,19)]
latest_basilicata <- basilicata[n,c(4,11,14,15,18,19)]
latest_calabria <- calabria[n,c(4,11,14,15,18,19)]
latest_campania <- campania[n,c(4,11,14,15,18,19)]
latest_e_romagna <- e_romagna[n,c(4,11,14,15,18,19)]
latest_fv_giulia <- fv_giulia[n,c(4,11,14,15,18,19)]
latest_lazio <- lazio[n,c(4,11,14,15,18,19)]
latest_liguria <- liguria[n,c(4,11,14,15,18,19)]
latest_lombardia <- lombardia[n,c(4,11,14,15,18,19)]
latest_marche <- marche[n,c(4,11,14,15,18,19)]
latest_molise <- molise[n,c(4,11,14,15,18,19)]
latest_bolzano <- bolzano[n,c(4,11,14,15,18,19)]
latest_trento <- trento[n,c(4,11,14,15,18,19)]
latest_piemonte <- piemonte[n,c(4,11,14,15,18,19)]
latest_puglia <- puglia[n,c(4,11,14,15,18,19)]
latest_sardegna <- sardegna[n,c(4,11,14,15,18,19)]
latest_sicilia <- sicilia[n,c(4,11,14,15,18,19)]
latest_toscana <- toscana[n,c(4,11,14,15,18,19)]
latest_umbria <- umbria[n,c(4,11,14,15,18,19)]
latest_v_aosta <- v_aosta[n,c(4,11,14,15,18,19)]
latest_veneto <- veneto[n,c(4,11,14,15,18,19)]
latest_italia <- italia[n,c(4,11,14,15,18,19)]
latest1 <- rbind(latest_abruzzo,latest_basilicata,latest_calabria,latest_campania,latest_e_romagna,
                 latest_fv_giulia,latest_lazio,latest_liguria,latest_lombardia,latest_marche,latest_molise,
                 latest_bolzano,latest_trento,latest_piemonte,latest_puglia,latest_sardegna,latest_sicilia,
                 latest_toscana,latest_umbria,latest_v_aosta,latest_veneto,latest_italia)
latest2 <- cbind(italia$date[n],latest1[order(latest1$totale_casi),])
names(latest2) <- c("Data","Regione","Totale positivi","Guariti","Deceduti","Totale casi","Tamponi effettuati")
print(latest2)
############################
############################
############################
```