---
title: "COVID-19"
author: "Maurizio La Rosa"
date: "20/10/2020"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Playing with COVID-19 data

This site is intended to let me work with Italian data on the COVID-19 infection spreading from the end of February, 2020 (and possibly, even before) and publish the results in a web site. 
While I have some previous experience working with R and R Markdown, it is new to me the application of Version Control techniques using Git (though I had a short approach with Git during my PhD years) and building a website.
So the results will be below any sufficient internet standard and, possibly, improving by the day (or week, or fill with any time unit at hand). 
The website is hosted in Github Pages and can be found here: <https://malrau.github.io/covid_19_data/>.

Here I will introduce the data When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r, echo=FALSE}
#setwd("github/covid_19_data/")
covid_data <- read.delim("covid_data.csv",sep=",",header=T)
covid_data$date <- as.Date(covid_data$date)
regions <- levels(covid_data$denominazione_regione)

```

Then I will build tables and plots that can be found in the appropriate tabs.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
